# Reproducible Research Assessment 1
====================================================


## Loading and preprocessing the data

```{r}
data <- read.csv( "activity.csv" )
summary( data )
```

## What is mean total number of steps taken per day?

### 1. Make a histogram of the total number of steps taken each day

Calculate total steps by date
```{r}
steps_by_date <- aggregate( steps ~ date, data = data, FUN = sum )
```

Make histogram
```{r}
hist( steps_by_date$steps )
```

### 2. Calculate and report the mean and median total number of steps taken per day

Calculate mean
```{r}
mean( steps_by_date$steps )
median( steps_by_date$steps )
```

## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Calculate mean of steps by interval
```{r}
mean_steps_by_interval <- aggregate( steps ~ interval, data = data, FUN = mean )
```

Make plot
```{r}
plot( mean_steps_by_interval$interval, mean_steps_by_interval$steps, type = "l" )
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

Find the maximum interval
```{r}
mean_steps_by_interval$interval[ which.max( mean_steps_by_interval$steps ) ]
```

## Imputing missing values

### 1. Calculate and report the total number of missing values in the dataset

```{r}
table( is.na( data$steps ) )[[2]]
```

### 2. Devise a strategy for filling in all of the missing values in the dataset

Use mean value of each interval to replace the missing values

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
no_missing_data <- data
for( i in 1 : nrow( data ) ){
    if( is.na( data$steps[ i ]  ) ){
      no_missing_data$steps[ i ] <- mean_steps_by_interval$steps[ mean_steps_by_interval$interval == data$interval[ i ] ]
    }
}
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

Make histogram of total steps by date
```{r}
no_missing_steps_by_date <- aggregate( steps ~ date, data = no_missing_data, FUN = sum )
hist( no_missing_steps_by_date$steps )
```

Mean and median of steps by date
```{r}
mean( no_missing_steps_by_date$steps )
median( no_missing_steps_by_date$steps )
```

Mean value doesn't change
Median value changes

## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r}
weekday <- weekdays( as.Date( no_missing_data$date ) ) == "Sunday" | weekdays( as.Date( no_missing_data$date ) ) == "Saturday"
no_missing_data$weekday <- factor( weekday,labels = c( "weekday", "weekend") )
```

### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend day

Create average steps by interval and weekday
```{r}
steps_by_interval_weekday <- aggregate( steps ~ interval + weekday, data = no_missing_data, FUN = mean )
```

Make plots
```{r}
library( lattice )
xyplot( steps ~ interval | weekday, data = steps_by_interval_weekday, type = "l", layout = c( 1, 2 ) )
```
